GPU-Beschleunigung der Material Point Method

### 2018-11-09 mscthesis.tex
singular value descr

\usepackage{listings}
\usepackage{xurl}
\urlstyle{sf}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{textcomp}
\usepackage{caption}
\usepackage{color}
\usepackage{subfigure}
\usepackage{wrapfig}
\usepackage[english,ngerman]{babel}
\usepackage[backend=biber,style=alphabetic]{biblatex}
\usepackage{algorithm}% http://ctan.org/pkg/algorithms
\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx
\addbibresource{mscthesis.bib}

@@

\cite{MPM:GPU}
%look again for MPM/PIC gpu impl
%SPH, FEM methods
\section{Notation}
Einstein-Notation
\section{Basics}
\subsection{Continuum Mechanics}
\cite{MPM:COURSE}
\subsection{Constitutive Models}
\cite{MPM:DRUCKER} %Sand Plasticity
\cite{MPM:MULTI} %Sand+Water
\cite{MPM:INVERT} %Neo-Hookean
\cite{MPM:OLROYDB} %Viscoelastic Foam,Sponge
\cite{MPM:SHELLS} %Thin Shells
\cite{MPM:PHASE_CHANGE} %Heat-Model
\cite{MPM:OPTIMI_INTEGR} %Opimization Methods, BDF-Semi-Impl-Midpoint rule

@@

  \boldsymbol{\epsilon} = \frac { 1 } { 2 } \left( \mathbf { F } + \mathbf { F } ^ { T } \right) - \mathbf { I }

@@

The energy produced by equation \ref{eq:linear_elas} using the small strain tensor is not rotationally invariant, w.r.t. to $\mathbf{F}$: $\psi(\mathbf{R}_{0} \mathbf{F}) \neq \psi(\mathbf{F})$. The deformation caused by stress however does not affect rigid body motions. Given the polar decomposition $\mathbf{F} = \mathbf{RS}$, an alternate strain measure may be defined as:

@@

  \hat { \boldsymbol { \epsilon } } ( \mathbf { F } ) = \boldsymbol { \epsilon } \left( \mathbf { R } ^ { T } \mathbf { F } \right) = \frac { 1 } { 2 } \left( \mathbf { R } ^ { T } \mathbf { F } + \left( \mathbf { R } ^ { T } \mathbf { F } \right) ^ { T } \right) - \mathbf { I } = \mathbf { S } - \mathbf { I }

@@

Substituting $\hat{\boldsymbol{\epsilon}}$ into equation \ref{eq:linear_elas} leads to the energy definition of corotational hyperelasticity using rotation-invariance of Frobenius norm and trace:

@@

  \Psi = \mu \| \mathbf { F } - \mathbf { R } \| _ { F } ^ { 2 } + \frac { \lambda } { 2 } \operatorname { tr } ^ { 2 } \left( \mathbf { R } ^ { T } \mathbf { F } - \mathbf { I } \right)

@@

\begin{equation}
  \Psi(\boldsymbol{\Sigma}) = \mu \| \boldsymbol { \Sigma } - \mathbf { I } \| _ { F } ^ { 2 } + \frac { \lambda } { 2 } \operatorname { tr } ^ { 2 } ( \boldsymbol { \Sigma } - \mathbf { I } )
\end{equation}
\cite{ADAMS:ELASTICITY}
\cite{MPM:INVERT}

\subsection{Singular Value Decomposition}
\cite{SVD:3x3}
\subsection{Material Point Method}
\subsubsection{Conservation Equations}
\cite{MPM:APIC}
\subsubsection{CFL Condition}
\subsection{GPU memory}
\subsubsection{Optimization opportunities}
\cite{NVIDIA:BEST:PRACTICE}
\cite{AMD:GPU_OPEN}
\begin{itemize}
  \item Minimize CPU-GPU transfers
  \item Coalesced memory access
  \item Bank conflicts
\end{itemize}
\subsubsection{SoA vs. AoS}
\section{Implementation}
\newcommand{\var}[1]{{\ttfamily#1}}% variable
\begin{algorithm}[t]
  \caption{Euclid’s algorithm}\label{euclid}
  \begin{algorithmic}[1]
    \Procedure{Euclid}{$a,b$}\Comment{The g.c.d. of a and b}
    \State $r\gets a\mathbfod b$ and \var{foobar}\label{foobar}
    \While{$r\not=0$}\Comment{We have the answer if r is 0}
    \State $a\gets b$
    \State $b\gets r$
    \State $r\gets a\mathbfod b$
    \EndWhile\label{euclidendwhile}
    \State \textbf{return} $b$\Comment{The gcd is b}
    \EndProcedure
  \end{algorithmic}
\end{algorithm}
\subsection{Framework}
decorator pattern
\subsection{Parallel Reduction}
\cite{NVIDIA:PARALLEL_REDUCTION}
\subsection{Binning}
\cite{NVIDIA:BINNING}
\subsection{Scanning}
\cite{NVIDIA:SCAN}
\cite{NVIDIA:SCAN_MODERN}
\subsection{Affine particle-in-cell}
\cite{MPM:APIC}
\subsection{Iterative solvers for linear systems}
%GMRES, Conjugate Gradients
\subsection{Verifying Simulation}
\begin{itemize}
  \item Conservation equations
  \item Singular Value Decomposition
  \item Parallel Reduction
\end{itemize}
\section{Evaluation}
\cite{NVIDIA:PEAK_PERFORMANCE}
\cite{KHRONOS:TIMER_QUERY}
\section{Conclusion}
\subsection{Review}
\subsection{Future Work}
Particle Activation: \cite{MPM:GPU}

>>>>

\usepackage{listings}
\usepackage{xurl}
\urlstyle{sf}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{textcomp}
\usepackage{caption}
\usepackage{color}
\usepackage{subfigure}
\usepackage{wrapfig}
\usepackage[english,ngerman]{babel}
\usepackage[backend=biber,style=alphabetic]{biblatex}
\usepackage{algorithm}% http://ctan.org/pkg/algorithms
\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx
\usepackage{mathtools}
\usepackage{subdepth}
\addbibresource{mscthesis.bib}

@@

\cite{MPM:GPU}
%look again for MPM/PIC gpu impl
%SPH, FEM methods
\section{Notation}
Einstein-Notation
\section{Basics}
\subsection{Continuum Mechanics}
\cite{MPM:COURSE}
\subsection{Singular Value Decomposition}
\cite{SVD:3x3}
\subsection{Constitutive Models}
\cite{MPM:DRUCKER} %Sand Plasticity
\cite{MPM:MULTI} %Sand+Water
\cite{MPM:INVERT} %Neo-Hookean
\cite{MPM:OLROYDB} %Viscoelastic Foam,Sponge
\cite{MPM:SHELLS} %Thin Shells
\cite{MPM:PHASE_CHANGE} %Heat-Model
\cite{MPM:OPTIMI_INTEGR} %Opimization Methods, BDF-Semi-Impl-Midpoint rule

@@

  \boldsymbol{\epsilon} = \frac { 1 } { 2 } \left( \mathbf { F }_E + \mathbf { F }_E ^ { T } \right) - \mathbf { I }

@@

The energy produced by equation \ref{eq:linear_elas} using the small strain tensor is not rotationally invariant, w.r.t. to $\mathbf{F}_E$: $\psi(\mathbf{R}_{0} \mathbf{F}_E) \neq \psi(\mathbf{F}_E)$. The deformation caused by stress however does not affect rigid body motions. Given the polar decomposition $\mathbf{F}_E = \mathbf{R}_E\mathbf{S}_E$, an alternate strain measure may be defined as:

@@

  \hat { \boldsymbol { \epsilon } } ( \mathbf { F }_E ) = \boldsymbol { \epsilon } \left( \mathbf { R }_E ^ { T } \mathbf { F }_E \right) = \frac { 1 } { 2 } \left( \mathbf { R }_E ^ { T } \mathbf { F }_E + \left( \mathbf { R }_E ^ { T } \mathbf { F }_E \right) ^ { T } \right) - \mathbf { I } = \mathbf { S }_E - \mathbf { I }

@@

Substituting $\hat{\boldsymbol{\epsilon}}$ into equation \ref{eq:linear_elas} leads to the energy definition of corotational hyperelasticity:
\begin{equation}\label{eq:corot_S}
  \Psi = \mu \| \mathbf{ S }_E - \mathbf { I } \| _ { F } ^ { 2 } + \frac { \lambda } { 2 } \operatorname { tr } ^ { 2 } ( \mathbf { S }_E - \mathbf { I } )
Using rotation-invariance of the Frobenius norm:

@@

  \Psi = \mu \| \mathbf {F}_E  - \mathbf {R}_E \| _ { F } ^ { 2 } + \frac { \lambda } { 2 } \operatorname { tr } ^ { 2 } \left( \mathbf { R }_E ^ { T } \mathbf { F }_E - \mathbf { I } \right)

@@

Even more insight yields the relationship to their singular values $\sigma _ i$, also called the principal stretches:
$$\operatorname {tr}\left(\mathbf{S}\right) = \sum _ { i = 1 } \sigma _ { i } = \operatorname{tr}\left(\mathbf{\Sigma}_E\right)$$
$$\| \mathbf{S}_E \| _ { F }^2 = \sum _ { i = 1 } \sigma _ { i } ^ { 2 } = \| \mathbf{\Sigma}_E\|_{F}^2$$
$$\Rightarrow \| \mathbf{S}_E\mathbf{-I} \| _ { F }^2 = \|\mathbf{S}_E \| _ { F }^2 - 2 \operatorname{tr}(\mathbf{S}_E)  + \|\mathbf{I}\|_{F}^2 = \| \boldsymbol { \Sigma }_E - \mathbf { I } \| _ { F } ^ { 2 }$$
\begin{equation} \label{eq:corot_sing}
  \Psi(\boldsymbol{\Sigma}_E) = \mu \| \boldsymbol { \Sigma }_E - \mathbf { I } \| _ { F } ^ { 2 } + \frac { \lambda } { 2 } \operatorname { tr } ^ { 2 } ( \boldsymbol { \Sigma }_E - \mathbf { I } )
Equation \ref{eq:corot_sing} uses the diagonal matrix $\mathbf{\Sigma}_E$ containing the principal stretches typically acquired by the singular value decomposition $\mathbf{F}_E \mathbf{= U \Sigma}_E\mathbf{ V ^T }.$ Since the energy density in \ref{eq:corot_sing} is a function of only three (singular) values, which describe stretch/compression of the material, isotropy of the material is underlined.
\cite{ADAMS:ELASTICITY}
Numerical Stepping algorithms
-> Newton Raphson scheme
-> Problem: Bad Primary contour
Benefit: Easier description in terms of Invariants
\cite{MPM:INVERT}

\subsection{Material Point Method}
\subsubsection{Conservation Equations}
\cite{MPM:APIC}
\subsubsection{CFL Condition}
\subsection{GPU memory}
\subsubsection{Optimization opportunities}
\cite{NVIDIA:BEST:PRACTICE}
\cite{AMD:GPU_OPEN}
\begin{itemize}
  \item Minimize CPU-GPU transfers
  \item Coalesced memory access
  \item Bank conflicts
\end{itemize}
\subsubsection{SoA vs. AoS}
\section{Implementation}
\newcommand{\var}[1]{{\ttfamily#1}}% variable
\begin{algorithm}[t]
  \caption{Euclid’s algorithm}\label{euclid}
  \begin{algorithmic}[1]
    \Procedure{Euclid}{$a,b$}\Comment{The g.c.d. of a and b}
    \State $r\gets a\mathbfod b$ and \var{foobar}\label{foobar}
    \While{$r\not=0$}\Comment{We have the answer if r is 0}
    \State $a\gets b$
    \State $b\gets r$
    \State $r\gets a\mathbfod b$
    \EndWhile\label{euclidendwhile}
    \State \textbf{return} $b$\Comment{The gcd is b}
    \EndProcedure
  \end{algorithmic}
\end{algorithm}
\subsection{Framework}
decorator pattern
\subsection{Parallel Reduction}
\cite{NVIDIA:PARALLEL_REDUCTION}
\subsection{Binning}
\cite{NVIDIA:BINNING}
\subsection{Scanning}
\cite{NVIDIA:SCAN}
\cite{NVIDIA:SCAN_MODERN}
\subsection{Affine particle-in-cell}
\cite{MPM:APIC}
\subsection{Iterative solvers for linear systems}
%GMRES, Conjugate Gradients
\subsection{Verifying Simulation}
\begin{itemize}
  \item Conservation equations
  \item Singular Value Decomposition
  \item Parallel Reduction
\end{itemize}
\section{Evaluation}
\cite{NVIDIA:PEAK_PERFORMANCE}
\cite{KHRONOS:TIMER_QUERY}
\section{Conclusion}
\subsection{Review}
\subsection{Future Work}
Particle Activation: \cite{MPM:GPU}
